\Large\textbf{Preparation GUIDE for Initial-Submission Manuscript}\\~\\ \large\textbf{ Preparation of Papers in Two-Column Format\\ for the ASP-DAC 2014 (\LaTeX2e version)}

### 2013-07-06 background.tex
Made the final changes

SystemJ consists of entities called clock-domains (CD)s, each running at
their individual pace, interacting with the environment via signals and
with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the schedule. We
call $\Delta$ the reaction time. Obviously, depending upon the amount of
computation required and schedule, $\Delta$ can vary, hence, in order to
satisfy the implicit restriction imposed by the synchrony hypothesis --
no input event \textit{from the environment} can be missed -- one needs
to calculate the \textit{Worst Case Reaction Time} (WCRT) and the
resultant WCRT needs to be smaller than the time between any two
consecutive events on any input, else the synchrony hypothesis is
violated. Techniques for calculating the WCRT of synchronous
programs~\cite{boldt07} exist and this analysis is not the focus of this
paper. The opposite of the WCRT is the \textit{Best Case Reaction Time}
(BCRT), we denote the WCRT and BCRT in Figure~\ref{fig:2c} using
\texttt{W} and \texttt{B}, respectively. Figure~\ref{fig:2c} shows a
continuously running physical clock (analogous to a clock in digital
hardware), the numerical annotations on the rising edge of the clock
mark the logical ticks from Figure~\ref{fig:2c}. The subscripts
\texttt{s} and \texttt{e} for the numbers show the start and end of the
logical ticks, respectively. The end of a logical tick and the start of
the next logical tick happen together. Figure~\ref{fig:2c} shows the
mapping of the logical time to the physical time. For example, the
logical tick \texttt{1} starts at the first rising edge and completes at
the second rising edge of the physical clock. Where as logical tick
\texttt{4} starts at the $4^{th}$ rising edge, but finishes at the
$6^{th}$ rising edge of the physical clock. This difference can be
accounted for by the java computations required when signal \texttt{A}
is input by the environment (see Figure~\ref{fig:2a}). This elasticity
is inherent (and elegant) to both Esterel and SystemJ.

>>>>

SystemJ consists of entities called \textit{Clock Domains} (CD)s, each
running at their individual pace, interacting with the environment via
signals and with each other via channels and CSP~\cite{choa85} style
rendezvous. Each CD has its own notion of logical time. CDs adheres to
the \textit{perfect synchrony} hypothesis, i.e., all statements execute
instantaneously in zero time. Only the \texttt{pause} statement consumes
time, just like in Esterel~\cite{gber931}. Consider the SystemJ program
in Figure~\ref{fig:2a}, it is waiting for an input signal producing
logical ticks (see Figure~\ref{fig:2c}). Once \texttt{A} is received at
tick 1 or 4, output \texttt{O} is emitted to the environment
instantaneously.

@@

The perfect synchrony hypothesis is ideal for programming the
synchronous sub-set of SystemJ. Unfortunately, every \textit{micro-step}
requires real-time to compute. Formally, execution of every micro-step
in the logical zero time, requires $\delta$ physical time. Time for a
logical tick can thus be summarized as $\Delta = \mathcal{F} (\delta)$,
where $\mathcal{F}$ is some function dependent upon the
\underline{schedule}. We call $\Delta$ the reaction time. Depending upon
the amount of computation required and schedule, $\Delta$ can vary,
hence, in order to satisfy the implicit restriction imposed by the
synchrony hypothesis -- no input event \textit{from the environment} can
be missed -- one needs to calculate the \textit{Worst Case Reaction
  Time} (WCRT) and the resultant WCRT needs to be smaller than the time
between any two consecutive events on any input, else the synchrony
hypothesis is violated. Techniques for calculating the WCRT of
synchronous programs~\cite{boldt07} exist and this analysis is not the
focus of this paper. The opposite of the WCRT is the \textit{Best Case
  Reaction Time} (BCRT), we denote the WCRT and BCRT in
Figure~\ref{fig:2c} using \texttt{W} and \texttt{B},
respectively. Figure~\ref{fig:2c} shows a continuously running physical
clock (analogous to a clock in digital hardware), the numerical
annotations on the rising edge of the clock mark the logical ticks from
Figure~\ref{fig:2c}. The subscripts \texttt{s} and \texttt{e} for the
numbers show the start and end of the logical ticks, respectively. The
end of a logical tick and the start of the next logical tick happen
together. Figure~\ref{fig:2c} shows the mapping of the logical time to
the physical time. For example, the logical tick \texttt{1} starts at
the first rising edge and completes at the second rising edge of the
physical clock. Where as logical tick \texttt{4} starts at the $4^{th}$
rising edge, but finishes at the $6^{th}$ rising edge of the physical
clock. This difference can be accounted for by the java computations
required when signal \texttt{A} is input by the environment (see
Figure~\ref{fig:2a}). This elasticity is inherent (and elegant) to both
Esterel and SystemJ.

### 2013-07-06 delays.tex
Made the final changes

We introduce a single \textit{derived} (consisting of kernel constructs)
statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
for real-time control. The resolution of the delay statement is of
secondary concern and is dependent upon the execution platform.

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a function $F$
maps these calculated values to a set of equidistant integer points
(values) separated by a unit value -- these points represent all the
logical ticks running at the WCRT and the BCRT, respectively that
satisfy the real-time delay requirements. The intersection of these two
sets gives all the logical ticks that satisfy the real-time requirements
invariant of the reaction time and its elasticity.

Let us revisit our motivating example to illustrate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50.3ms and $N$ is 200.3ms,
respectively. Let the WCRT and BCRT be: 0.112ms and 0.0334ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can pick up any value
  for \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantee while preserving the essence of synchronous and
GALS programming prescribed by SystemJ and Esterel style
languages. Moreover, this technique considers non maximal-parallelism,
i.e., the delays in logical ticks is calculated after scheduling of CDs
or synchronous programs has been performed. To our knowledge we are the
first to do so.

@@

The relaxation algorithm is shown in Algorithm~\ref{alg:2}.

@@

Many real-world systems require non-deterministic timing constructs. One
of them is our motivating example -- the human response time
system.

@@

% \begin{verbatim}
% {await (job); delay(2..10 ms);  //SPOOLER CD
%  send mid(job) delay (1ms);}
% ><
% {receive mid; delay(1..30 ms); //PRINTER CD
%  emit print(#mid);}
% \end{verbatim}
% \end{scriptsize}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||{delay((1..1)ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

@@

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10ms to pass by. Consider what happens when
signals \texttt{S} and \texttt{TIMER} occur in the same logical tick,
the \texttt{await} statement is never executed and hence, we enter a
deadlock. Such problems are completely avoided in our technique, because
we convert the real-time delays into logical delays (\texttt{pause}
constructs), which interact well with preemption.

@@

Channels, used for communication between reactions in asynchronously
running CDs, are an addition in the SystemJ language. Like interaction
of preemption and delays, conversion of rela-time delays to logical
ticks also interacts well with channel rendezvous, because the semantics
of interaction are well defined~\cite{amal10}. More importantly, we need
to consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. % The response time to input signals \textit{is} though!
% But, we are unconcerned with the response time, analysis in this paper
% and it remains a future research avenue.

>>>>

We introduce a single \textit{derived} (built from kernel constructs)
statement called \mbox{\texttt{delay (M..N)}} in the SystemJ language
for real-time control. The resolution of the delay statement is of
secondary concern and is dependent upon the execution platform.

@@

Our algorithm takes as input: WCRT, BCRT, and the lower and upper bounds
$M$ and $N$, respectively of the \texttt{delay} construct. We first
divide $M$ and $N$ with the BCRT and WCRT, respectively. This division
gives us the number of individual logical ticks required to delay the CD
(or synchronous program) by the real-time specification. We always
$ceil$ when dividing $M$ and $floor$ when dividing $N$ to make sure that
the resultant values are integers (in domain $\mathbb{N}^{>0}$) and
these functions guarantee that the resultant logical ticks result in
real-time delays between the required range $M-N$. Next, a function $F$
maps these calculated values to a set of equidistant integer points
(values) separated by a unit value -- these points represent all the
logical ticks running at the WCRT and the BCRT, respectively that
satisfy the real-time delay requirements. The intersection of these two
sets gives all the logical ticks that satisfy the real-time requirements
invariant of the logical time and its elasticity.

Let us revisit our motivating example to illustrate the algorithm. From
Figure~\ref{fig:1} we know that $M$ is 50.3 ms and $N$ is 200.3 ms,
respectively. Let the WCRT and BCRT be: 0.112 ms and 0.0334 ms,
respectively. Thus, the algorithm proceeds as follows:

\begin{enumerate*}
\item $l_1 \leftarrow \lceil 50.3/0.112 \rceil$ and $u_1 \leftarrow
  \lfloor 200.3/0.112 \rfloor$. $l_1 = 450$ and $u_1 = 1788$. We first
  calculate the logical ticks that are always running at the WCRT and
  satisfy the required real-time delay.
\item $l_2 \leftarrow \lceil 50.3/0.0334 \rceil$ and $u_2 \leftarrow
  \lfloor 200.3/0.0334 \rfloor$. $l_1 = 1506$ and $u_1 = 5997$. We do
  the same for the BCRT case.
\item $S_1 = [450..1788]$ and $S_2 =[1506,5997]$. We then map the
  resultant bounds to linear points. Sets $S_1$ and $S_2$ represent
  logical ticks that running at the WCRT and BCRT, respectively always
  satisfy the required real-time constraints.
\item $D = S_1 \cap S_2$, $D = [1506,1788]$. Finally, the intersection
  of the two sets gives the set $D$ from which we can choose any value
  for \texttt{d}.
\end{enumerate*}

The resultant value for \texttt{d} gives the number of logical ticks,
which can run at any physical clock-speed, bounded by the BCRT and the
WCRT of the program and still result in the desired real-time delay. We
think this is an elegant solution, because the technique provides
hard-real time guarantees while preserving the essence (elastic logical
tick) of synchronous and GALS programming prescribed by SystemJ and
Esterel style languages. Moreover, this technique considers non
maximal-parallelism, i.e., the delays in logical ticks is calculated
after scheduling of CDs or synchronous programs has been performed. To
our knowledge we are the first to do so.

@@

The relaxation technique is shown in Algorithm~\ref{alg:2}.

@@

Many real-world systems require non-deterministic timing
constructs. Where the exact real-time delay is not known or should not
be known apriori. One such system is our motivating example -- the human
response time system.

@@

% \begin{verbatim}
% {await (job); delay (2..10 ms);  //SPOOLER CD
%  send mid(job) delay (1ms);}
% ><
% {receive mid; delay(1..30 ms); //PRINTER CD
%  emit print(#mid);}
% \end{verbatim}
% \end{scriptsize}

@@

There are many instances when one would like to wait on an input from
environment for only a specified amount of time. This can be programmed
as:
% \begin{scriptsize}
\begin{verbatim}
// timeout after 1ms.
trap(T){{await(A);}||{delay(1..1 ms);exit(T);}}
\end{verbatim}
% \end{scriptsize}

@@

As identified in~\cite{Bourke2009a} the \texttt{suspend} does not play
well with the external timer. The above program sends a signal to an
external timer and waits for 10 ms to pass by. Consider what happens
when signals \texttt{S} and \texttt{TIMER} occur in the same logical
tick, the \texttt{await} statement is never executed (due to suspend)
and hence, we enter a deadlock. Such problems are completely avoided in
our technique, because we convert the real-time delays into logical
delays (\texttt{pause} constructs), which interact well with preemption.

@@

Channels, used for communication between reactions in asynchronously
running CDs, are an addition in the SystemJ language. Like interaction
of preemption and delays, conversion of real-time delays to logical
ticks also interacts well with channel rendezvous, because the semantics
of interaction are well defined~\cite{amal10}. More importantly, we need
to consider the interplay of channel communication and WCRT/BCRT
analysis. Since, channel communication does not stop logical time
(see~\cite{amal10}) WCRT/BCRT are unaffected by channel
communication. % The response time to input signals \textit{is} though!
% But, we are unconcerned with the response time, analysis in this paper
% and it remains a future research avenue.

### 2013-07-06 introduction.tex
Made the final changes

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc.  Usually these systems also need to meet
real-time constraints enforced by the environment.  Yet, the languages
used to program these systems do not support describing real-time
constructs as first class language constructs.  For example, one cannot
describe a simple real-time delay (postponement) of an operation for 0.2
ms. These languages are based on formal semantics, essential for the
formal reasoning about and verification of correctness of functional
properties of the developed programs, but leave nonfunctional properties
such as timing behavior as an implementation
detail~\cite{boldt07}. Arguably, rightly so, because physical time
cannot be incorporated without information about the underlying
execution platform.  But, one can still reason in terms of logical
time. These languages support discrete logic clock rather than a
discrete physical (real-time) clock. The real-time period of the
discrete logic clock (also referred to as a logic tick) is not fixed and
it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logic clock is elastic. Although the timing
model with logic clock has worked well for reactive synchronous
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as the ﬁrst class language
constructs create additional opportunity for formally verifying
functional and non-functional requirements before system deployment.

@@

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delays model relays on
the use of relative instead of absolute time, i.e., a delay in selected
time units is counted from the currently executing SystemJ statement and
(3) delays allow certain behaviours to having real-time features, while
the other behaviours use only logical time and (4) real-time delay does
not affect the model of logical time in SystemJ program as long as the
amount of delay is within boundaries that can be determined statically
by program analysis. All these as well as semantics of the delay
construct are discussed and illustrated in the remaining part of the
paper. Our contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reduced power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify a
  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., Figure~\ref{fig:1}.
\end{enumerate*}

>>>>

Reactive languages~\cite{gber931,amal10} are a class of programming
languages used for designing and implementing reactive systems, which
continuously respond to input from their environment. These languages
have been successfully used in programming a plethora of systems such as
fly-by-wire in Airbus~\cite{eairbus}, security surveillance
systems~\cite{amal121}, etc.  Usually these systems also need to meet
real-time constraints enforced by the environment. Yet, these languages
do not support describing real-time statements as first class language
constructs.  For example, one cannot describe a simple real-time delay
(postponement) of an operation for 0.2 ms. These languages are based on
formal semantics, essential for the formal reasoning about and
verification of correctness of functional properties of the developed
programs, but leave nonfunctional properties such as timing behavior as
an implementation detail~\cite{boldt07}. Arguably, rightly so, because
physical time cannot be incorporated without information about the
underlying execution platform.  But, one can still reason in terms of
logical time. These languages support discrete logical clock rather than
a discrete physical (real-time) clock. The real-time period of the
discrete logical clock (also referred to as a logical tick) is not fixed
and it is determined by the responsiveness of the program to the input
signals. Unlike a discrete physical clock, which has fixed real-time
period, the period of the logical clock is elastic. Although the timing
model with logical clock has worked well for reactive synchronous/GALS
languages in designing discrete systems, there is a need to introduce
real-time since, often, implementation models are in the continuous or
discrete real-time domain~\cite{DBLP:journals/pieee/SifakisTY03}.
Moreover, adding timing capabilities as first class language constructs
create additional opportunity for formally verifying functional and
non-functional requirements before system deployment.

@@

The major motivations for introducing delay mechanism and at the same
time contributions of the paper are: (1) delays allow real-time based
synchronization between concurrent behaviors, (2) delays model relays on
the use of relative instead of absolute time, i.e., a delay in selected
time units is counted from the currently executing SystemJ statement and
(3) delays allow mixing behaviors with real-time features and others
with only logical time and (4) real-time delay does not affect the model
of logical time in SystemJ program as long as the amount of delay is
within boundaries that can be determined statically by program
analysis. All these as well as semantics of the delay construct are
discussed and illustrated in the remaining part of the paper. Our
contributions can be further refined as follows:
\begin{enumerate*}
\item Real-time is modeled in the $\mathbb{Q}^{>0}$ domain of numbers,
  i.e., non-negative rational numbers.
\item We consider the property of \textit{non
    maximal-parallelism}. \textit{Non-maximal parallelism:} There are
  not always sufficient resources for processes to execute, so
  effectively processes are allocated and scheduled with different
  optimization criteria -- reducing computation time, reducing power
  consumption, etc. This allocation and scheduling is an integral part
  of any language compilation and needs to be considered when
  introducing real-time delays.
% \item Our solution does not change the functional or timing semantics of
%   GALS and its subset synchronous programs, in fact, our solution does
%   not require one to change the mid-end, the back-end or the related
%   optimization phases of the compiler at all.
\item We are the first, to our knowledge, to allow specification of both
  deterministic and non-deterministic real-time in GALS and synchronous
  languages. \textit{Deterministic delay:} is the ability to specify a
  single \textit{real-time} delay. \textit{Non-deterministic delay:} is
  the ability to specify a range, of real-time values for a delay
  statement, e.g., delay statement in Figure~\ref{fig:1}.
\end{enumerate*}

### 2013-07-06 motivation.tex
Made the final changes

\begin{verbatim}
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
    //exit after any 
    // where from 50.3 to 200.3 ms.
    delay (50.3 .. 200.3 ms);
    exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S;}
  // do something with S
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a simple machine used to test human response time and collect
this data for future research purposes. The machine switches on a green
light on a touch screen for anywhere between \texttt{50.3ms} to
\texttt{200.3ms}, if the user can touch this green light then she is
successful and an integer \texttt{1} is sent to a database, if
unsuccessful a \texttt{0} is sent instead, Different statistics and
analyses on human responsiveness can be based on collected data.

The SystemJ psuedocode~\cite{amal10} code for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of
\texttt{50.3ms} to \texttt{200.3ms} a positive response is sent to the
second clock-domain. All the SystemJ programming constructs used to
implement this system are described in~\cite{amal10} and
Table~\ref{tab:1} in Section~\ref{sec:background}. The delay
speciﬁcation, which models the non-deterministic delay between 50.3ms
and \texttt{200.3ms} is not supported in the current language and cannot
be compiled with the current SystemJ compiler.

>>>>

\begin{verbatim}
{ // Clock-domain 1
while(true) {
  trap(TRAP){
   // Abort if user touches the screen
   abort(TOUCH){
    {sustain GREEN_LIGHT;}
    || // synchronous parallel operator
    {
     //exit after any where from 50.3 to 200.3 ms.
     delay (50.3 .. 200.3 ms);
     exit(TRAP); 
    }
   } do {send S(1);} // Send 1 if touched 
  } do {send S(0);} // Else Send 0
  pause; // only construct to consume time
 }
}
>< // asynchronous operator
{ // Clock-domain 2
  while(true){ receive S; // do something with S}
}
\end{verbatim}
  \end{scriptsize}
\end{minipage}
\caption{An example human responsiveness system (HRTCS)}
% }
% \end{center}
\label{fig:1}
\end{figure}

Consider a simple machine used to test human response time and collect
this data for future research purposes. The machine switches on a green
light on a touch screen for anywhere between 50.3 ms to 200.3 ms, if the
user can touch this green light then she is successful and an integer
\texttt{1} is sent to a database, if unsuccessful a \texttt{0} is sent
instead, Different statistics and analyses on human responsiveness can
be based on collected data.

The SystemJ psuedocode~\cite{amal10} for such a system is shown in
Figure~\ref{fig:1}. We use the SystemJ language in this paper, first of
all, because being a \textit{Globally Asynchronous Locally Synchronous}
(GALS) language it helps us easily describe a larger class of systems
(like HRTCS in Figure~\ref{fig:1}) compared to purely synchronous
languages like Esterel. Moreover, being a super-set of Esterel, SystemJ
allows us to explore the marriage of real-time and logical time, which
is equally applicable to the synchronous sub-set.

The SystemJ program is divided into two clock-domains, the first
clock-domain continuously displays a green light (\texttt{GREEN\_LIGHT}
signal) on the touch screen sensor and waits for the user to respond. If
the user is able to touch the screen within the specified time of 50.3
ms to 200.3 ms a positive response is sent to the second
clock-domain. All the SystemJ programming constructs used to implement
this system are described in~\cite{amal10} and Table~\ref{tab:1} in
Section~\ref{sec:background}. The delay speciﬁcation, which models the
non-deterministic delay between 50.3 ms and 200.3 ms is not supported in
the current language and cannot be compiled with the current SystemJ
compiler. We introduce real-time \texttt{delay} statements in SystemJ.
